---
title: "p8105_hw1_fl2569"
author: "Fang Liu"
date: "9/27/2021"
output: github_document
---

### Problem 1

This problem is intended to emphasize variable types and introduce coercion; it is important to know how R treats numeric, character, and factor variables.

First, load in the required tidyverse library.
```{r setup, include = FALSE}
library(tidyverse)
```

Then, create the data frame containing the specified variables. 
```{r df_create}
set.seed(1234)

q1_dataframe <- tibble(
  norm_samp = rnorm(10),
  norm_samp_pos = norm_samp > 0,
  vec_character = c("hello","my", "name", "is", "Amy", "how", "are", "you", "doing", "today"),
  vec_factor = factor(c("Small", "Medium", "Large", "Small", "Medium", "Large","Small", "Small", "Medium", "Large"))
)
```

Take the mean of each variable in the data frame. 
```{r calc_mean, eval = FALSE}
mean_samp <- mean(q1_dataframe$norm_samp)
mean_pos <- mean(q1_dataframe$norm_samp_pos) #returns the probability of TRUE
mean_vector <- mean(q1_dataframe$vec_character) #Warning message: argument is not numeric or logical: returning NA
mean_factor <- mean(q1_dataframe$vec_factor) #Warning message: argument is not numeric or logical: returning NA
```

We can take the mean of the generated normal numbers; the mean of the logic vector give us the proportion of the numbers that are positive. However, we can't take the mean of a vector or factor so NA is returned. 