---
title: "p8105_hw1_fl2569"
author: "Fang Liu"
date: "9/27/2021"
output: github_document
---

### Problem 1

This problem is intended to emphasize variable types and introduce coercion; it is important to know how R treats numeric, character, and factor variables.

First, load in the required tidyverse library.
```{r setup, message = FALSE}
library(tidyverse)
```

Then, create the data frame containing the specified variables. 
```{r df_create}
set.seed(1234)

q1_dataframe <- tibble(
  norm_samp = rnorm(10),
  norm_samp_pos = norm_samp > 0,
  vec_character = c("hello","my", "name", "is", "Amy", "how", "are", "you", "doing", "today"),
  vec_factor = factor(c("Small", "Medium", "Large", "Small", "Medium", "Large","Small", "Small", "Medium", "Large"))
)
```

Take the mean of each variable in the data frame. 
```{r calc_mean, eval = FALSE}
mean_samp <- mean(q1_dataframe$norm_samp)
mean_pos <- mean(q1_dataframe$norm_samp_pos) #returns the probability of TRUE
mean_vector <- mean(q1_dataframe$vec_character) #Warning message: argument is not numeric or logical: returning NA
mean_factor <- mean(q1_dataframe$vec_factor) #Warning message: argument is not numeric or logical: returning NA
```

We can take the mean of the generated normal numbers; the mean of the logic vector give us the proportion of the numbers that are positive. However, we can't take the mean of a vector or factor so NA is returned. 

Next, try use the **as.numeric** function to convert the logical, character, and factor variables into numericals.
```{r as.numeric, eval = FALSE}
coercion_logical <- as.numeric(q1_dataframe$norm_samp_pos)
coercion_character <- as.numeric(q1_dataframe$vec_character) #Warning message:NAs introduced by coercion 
coercion_factor <- as.numeric(q1_dataframe$vec_factor) #gives us which factor level each variable belongs to
```

When we try to convert character variables into numeric, we get a warning message and all values are converted to NA; thus, we couldn't take the mean. However, when we convert the factor variables into numerical, we get the factor level each variable belongs to. 


### Problem 2

This problem focuses the use of inline R code, plotting, and the behavior of ggplot for variables of different types.Remember to install the palmerpenguins package in the console! Then, load in the penguins data set. 
```{r load_penguins, message = FALSE}
data("penguins", package = "palmerpenguins")
```

#### Describe the dataset
The penguins data set has `r nrow(penguins)` rows and `r ncol(penguins)` columns. The data set contains information about what species the penguins belong to (i.e., Adelie, Chinstrap, or Gentoo), what island they live on (i.e., Biscoe, Dream, or Torgersen), their bill length and depth in millimeter, the flipper length, their body mass, gender, and the year. The mean flipper length is `r mean(pull(penguins, flipper_length_mm))`.

#### Create scatterplot
Make a scatter plot of flipper length vs. bill length and color the points using the species variable. Lastly, export the scatterplot to the project directory using ggsave .
```{r flipper_bill_scatter, warning = FALSE}
ggplot(penguins, aes(x=bill_length_mm, y = flipper_length_mm, color=species)) + geom_point()
ggsave("scatter_plot.pdf", height = 4, width = 6)
```
